Illustrating Good Practice in {\LaTeX}

### 2021-09-06 articles/good_practice.tex
added header and footer
fixed order of sections
minor change
keeping things up to date
twenty third payload
major improvement and fixes

\title{Illustrating Good Practice in {\LaTeX}}
\author{Nicholas D. Kostin}
\date{July 2021}

@@

\begin{table}[H]
\centering
\captionsetup{width=0.8\textwidth,labelfont={color=black,bf},textfont={color=black}}
\caption*{Maxwell's equations and their interpretation.}
\begin{tabular}{@{}c|c@{}}
\toprule
Equation & Interpretation \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{gather*} \nabla \cdot \bm{E} = \frac{\rho}{\epsilon_o} \end{gather*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{E}$-fields with divergence come from charges (electric monopoles). \end{center}}} \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{gather*} \nabla \cdot \bm{B} = 0 \end{gather*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{B}$-fields with divergence don't exist (no magnetic monopoles). \end{center}}} \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{gather*} \nabla \times \bm{E} = -\frac{\partial \bm{B}}{\partial t} \end{gather*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{E}$-fields with curl come from time-varying $\bm{B}$-fields --- and those \emph{only}. \end{center}}} \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{align*} \nabla \times \bm{B} &= \mu_o \bm{J} + \mu_o \epsilon_o \frac{\partial \bm{E}}{\partial t} \\ &= \mu_o \left( \bm{J} + \bm{J}_D \right) \end{align*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{B}$-fields with curl come from currents (moving electric monopoles) and from time-varying $\bm{E}$-fields. Sometimes we refer to $\epsilon_o \frac{\partial \bm{E}}{\partial t}$ as displacement current and bundle the $J$'s together. \end{center}}} \\ \bottomrule
\end{tabular}
\end{table}

@@

Shown on the following pages is the implementation of this numerical method.

@@

>>>>

% Continuous per-section numbering of figures and tables
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}

\title{Illustrating Good Practice in {\LaTeX}}
\author{Nicholas D. Kostin}
\date{October 2021}

\usepackage{fancyhdr,lastpage}
\pagestyle{fancy}
\fancyhf{}
\lhead{Illustrating Good Practice in {\LaTeX}}
\rhead{Nicholas D. Kostin}
\lfoot{\texttt{https://nkostin.xyz}}
\rfoot{\thepage\ of \pageref{LastPage}}

@@

\begin{table}[H]
\centering
\captionsetup{width=0.8\textwidth,labelfont={color=black,bf},textfont={color=black}}
\caption{Maxwell's equations and their interpretation.}
\begin{tabular}{@{}c|c@{}}
\toprule
Equation & Interpretation \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{gather*} \nabla \cdot \bm{E} = \frac{\rho}{\epsilon_o} \end{gather*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{E}$-fields with divergence come from charges (electric monopoles). \end{center}}} \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{gather*} \nabla \cdot \bm{B} = 0 \end{gather*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{B}$-fields with divergence don't exist (no magnetic monopoles). \end{center}}} \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{gather*} \nabla \times \bm{E} = -\frac{\partial \bm{B}}{\partial t} \end{gather*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{E}$-fields with curl come from time-varying $\bm{B}$-fields --- and those \emph{only}. \end{center}}} \\ \midrule
{\parbox[c]{0.3\linewidth}{\begin{align*} \nabla \times \bm{B} &= \mu_o \bm{J} + \mu_o \epsilon_o \frac{\partial \bm{E}}{\partial t} \\ &= \mu_o \left( \bm{J} + \bm{J}_D \right) \end{align*}}} & {\parbox[c]{0.6\linewidth}{\begin{center} $\bm{B}$-fields with curl come from currents (moving electric monopoles) and from time-varying $\bm{E}$-fields. Sometimes we refer to $\epsilon_o \frac{\partial \bm{E}}{\partial t}$ as displacement current and bundle the $J$'s together. \end{center}}} \\ \bottomrule
\end{tabular}
\end{table}

@@

What follows is the implementation of this numerical method.

@@

Don't forget, however, that too much color is cringe, so be exercise caution!

\section{A Particle in 3D}

A particle in 3D has mass $m$ and is in the potential $V(x,y,z) = \frac{1}{2} m \omega^2 \left( x^2 + y^2 + z^2 \right).$

\subsection{Energies of the Lowest Four Eigenstates}

Suppose we need to find the energies of the four lowest eigenstates of the Hamiltonian and their corresponding degeneracies. We can start by expanding the potential as
\begin{equation*}
    V(x, y, z) = \underbrace{\frac{1}{2} m \omega^2 x^2}_{V_x(x)} + \underbrace{\frac{1}{2} m \omega^2 y^2}_{V_y(y)} + \underbrace{\frac{1}{2} m \omega^2 z^2}_{V_z(z)}.
\end{equation*}

And the total eigenenergy is
\begin{equation}
    E_{n_x n_y n_z} = E_{n_x} + E_{n_y} + E_{n_z} = \hbar \omega \left( n_x + \frac{1}{2} \right) + \hbar \omega \left( n_y + \frac{1}{2} \right) + \hbar \omega \left( n_z + \frac{1}{2} \right) \label{eq:3dQHO:eigenenergy}
\end{equation}

where $n_x, n_y, n_z = 0, 1, 2, \dots$ are independent from each other. The ground state should have $n_x = n_y = n_z = 0$, so its energy is given by

\begin{equation*}
    E_{000} = 3 \hbar \omega \left( \frac{1}{2} \right) = \frac{3}{2} \hbar \omega.
\end{equation*}

The first excited state has energy
\begin{equation*}
    \underbrace{E_{100} = E_{010} = E_{001}}_{\text{degeneracy 3}} = \frac{\hbar}{\omega} \left( \frac{3}{2} + \frac{1}{2} + \frac{1}{2} \right) = \frac{5}{2} \hbar \omega.
\end{equation*}

The second excited state has energy
\begin{equation*}
    \underbrace{E_{200} = E_{020} = E_{002} = E_{110} = E_{101} = E_{011}}_{\text{degeneracy 6}} = \frac{\hbar}{\omega} \left( \frac{5}{2} + \frac{1}{2} + \frac{1}{2} \right) = \frac{7}{2} \hbar \omega.
\end{equation*}

And finally, the third excited state has energy
\begin{equation*}
    \underbrace{\left\{ \begin{array}{l} E_{300} = E_{030} = E_{003} = E_{111} = E_{102} \\[0.4em] E_{120} = E_{201} = E_{021} = E_{210} = E_{012} \end{array} \right\}}_{\text{degeneracy 10}} = \frac{\hbar}{\omega} \left( \frac{7}{2} + \frac{1}{2} + \frac{1}{2} \right) = \frac{9}{2} \hbar \omega.
\end{equation*}

These results are summarized in the table below.

\begin{table}[H]
\centering
\captionsetup{width=0.8\textwidth,labelfont={color=BLACK,bf},textfont={color=BLACK}}
\caption{Energies and corresponding degeneracies of the 3D QHO}
\begin{tabular}{@{}ccc@{}}
\toprule
State & Energy & Degeneracy \\ \midrule
\multicolumn{1}{c|}{Ground state} & \multicolumn{1}{c|}{$\displaystyle \frac{3}{2}\hbar\omega$} & 1 \\[1em]
\multicolumn{1}{c|}{First excited state} & \multicolumn{1}{c|}{$\displaystyle \frac{5}{2}\hbar\omega$} & 3 \\[1em]
\multicolumn{1}{c|}{Second excited state} & \multicolumn{1}{c|}{$\displaystyle \frac{7}{2}\hbar\omega$} & 6 \\[1em]
\multicolumn{1}{c|}{Third excited state} & \multicolumn{1}{c|}{$\displaystyle \frac{9}{2}\hbar\omega$} & 10 \\ \bottomrule
\end{tabular}
\end{table}

\subsection{Degeneracy of an Arbitrary Excited State}

Now suppose we need to find the degeneracy $g_n$ of the $n^{th}$ excited state as a function of $n$. For our three-dimensional harmonic oscillator, we can define
\begin{equation*}
    n = n_x + n_y + n_z.
\end{equation*}

It follows from Eq. (\ref{eq:3dQHO:eigenenergy}) that the energy is strictly a function of $n$. That is, any combination of $n_x$, $n_y$, and $n_z$, as long as they produce the same $n$, will yield the same energy. Such states are said to be \emph{degenerate}. Our goal is to find how many possible combinations of $n_x$, $n_y$, and $n_z$ exist for a given $n$.

To explicitly determine the degeneracy $g_n$ for a given state, we fix the value of $n$. Because of the restriction imposed by setting $n = n_x + n_y + n_z$, we really have two free parameters (say, $n_x$ and $n_y$). We let $n_x$ assume any allowable integer value:
\begin{equation*}
    n_x \in \left\{ 0, 1, \dots, n \right\}.
\end{equation*}

Then $n_y$ can assume any integer value given the sum $n_x + n_y$ does not exceed $n$:
\begin{equation*}
    n_y \in \left\{ 0, 1, \dots, n - n_x \right\}.
\end{equation*}

In particular, for any value of $n_x$, there are $n - n_x + 1$ unique choices for $n_y$ (note the size of the set of permissible $n_y$ values). Finally, $n_z$ is automatically fixed:
\begin{equation*}
    n_z = n - n_x - n_y.
\end{equation*}

For each permissible value of $n_x$ (from $0$ to $n$), there are $n - n_x + 1$ choices for $n_y$. Summing the number of these choices for all allowable values of $n_x$ yields
\begin{align*}
    \sum\limits_{n_x = 0}^n \left( n - n_x + 1 \right) &= \left(n + 1\right) + n + \left(n - 2\right) + \cdots + \cancelto{0}{\left(n - n\right)} \\
                                                       &= \frac{\left(n+1\right) \left(n+2\right)}{2}.
\end{align*}

The series above is well known. Thus, we find that the degeneracy is
\begin{equation*}
    g_n = \frac{\left(n+1\right) \left(n+2\right)}{2}.
\end{equation*}
